---
title: "alpha analysis"
author: "Elizabeth Hiroyasu"
date: "March 8, 2018"
output: html_document
---
In order to estimate the alpha parameter for the model, we simulate the model over different values of alpha to see which yields a type II functional response

```{r setup}
setwd("../GophersVolesOwls")
library(devtools)
library(roxygen2)
load_all("owls")
library(owls)
```

Setting up the state variables and varying the alpha parameter
```{r}
#State variables
state<-c(N = 87.5, P = 0.6)
times<- seq(0, 20, by = 1)

#Parameters
alpha<-seq(0.01, 10, by=0.5)
beta = 1.01e-3
delta=0.01
K_prey=175
k_max=378
h=1/k_max
D=1/(alpha*h)
r=0.28

parameters<-list(length(alpha))
for (i in 1:length(alpha)){
  parameters[[i]] <- c(r, K_prey, k_max, D[i])
}
```

Looping over these values
```{r}
out_alpha_sim<-vector("list", length(alpha))
names(out_alpha_sim)=alpha

for(i in 1:length(parameters)){
  out_alpha_sim[[i]]<- ode(y=state, times = times, func = pred_prey, parms = c(r=r, K_prey=K_prey, k_max=k_max, D=D[i]))
}
```
note that D is the term that is varying here because D=1/alpha*h

Plotting N versus time, grouped by alpha
```{r}
alpha_sim<-ldply(out_alpha_sim, data.frame)
NvT<-ggplot(data=alpha_sim[,-4], aes(x=time, y=N, group=as.factor(.id), color=as.factor(.id)))+geom_line() +
  labs(title="N vs time", x="time", y="N") + scale_color_discrete(name="alpha")
NvT
```

Plotting alpha value vs. N at time 10
```{r, echo=FALSE}
NAlpha<-subset(alpha_sim, time==max(times))
ggplot(data=NAlpha[,-4], aes(x=.id, y=N)) + geom_point() + labs(title="alpha vs N at t final", x="alpha", y="N")
```

This suggests that coexistence can occur between alpha=0.01 and alpha=1.0

