---
title: "Prey Control"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
##Extinction of gopher populations
```{r setup, message=FALSE}
library(tidyverse)
gopher_df<-data.frame(read_csv('./Data/processed/GopherTable042318.csv'))
vole_df<-data.frame(read_csv('./Data/processed/VoleTable042318.csv'))
```


Suppose we're interested in how often the prey populations drop below two individuals
```{r}
count(gopher_df, N<2)
count(vole_df, N<2)
```
Across all possible scenarios, the population of gophers is reduced to less then two individuals 134 times out of 243 scenario result - over 50% of the time. The population of voles is reduced to less than two individuals 177 times out of 243 resulting in extinction over 70% of the time.

For gophers we can investigate the relationship between alpha, r, and prey extinction.
```{r}
gopher_sum_alpha<- gopher_df %>%
    group_by(alpha, r) %>%
    mutate(ext = N<2) %>%
    summarise(mean_gopher=mean(ext))

ggplot(gopher_sum_alpha, aes(as.factor(alpha), mean_gopher))+
  geom_bar(stat='identity')+
  facet_grid(.~r) +
  labs(title="Percentage of simulations when gophers go extinct", 
       x="Alpha", y="Percent of simulation runs N<2 individuals")
```

We can also investigate the relationship between predator density, r, and prey extinction.
```{r}
gopher_sum_P<- gopher_df %>%
    group_by(P, r) %>%
    mutate(ext = N<2) %>%
    summarise(mean_gopher=mean(ext))

ggplot(gopher_sum_P, aes(as.factor(P), mean_gopher))+
  geom_bar(stat='identity')+
  facet_grid(.~r) +
  labs(title="Percentage of simulations when gophers go extinct", 
       x="Predator Density", y="Percent of simulation runs N<2 individuals")
```

We can do the same for voles.
```{r}
vole_sum_alpha<- vole_df %>%
    group_by(alpha, r) %>%
    mutate(ext = N<2) %>%
    summarise(mean_vole=mean(ext))

ggplot(vole_sum_alpha, aes(as.factor(alpha), mean_vole))+
  geom_bar(stat='identity')+
  facet_grid(.~r) +
  labs(title="Percentage of simulations when voles go extinct", 
       x="Alpha", y="Percent of simulation runs N<2 individuals")

vole_sum_P<- vole_df %>%
    group_by(P, r) %>%
    mutate(ext = N<2) %>%
    summarise(mean_vole=mean(ext))

ggplot(vole_sum_P, aes(as.factor(P), mean_vole))+
  geom_bar(stat='identity')+
  facet_grid(.~r) +
  labs(title="Percentage of simulations when voles go extinct", 
       x="Predator Density", y="Percent of simulation runs N<2 individuals")
```



