---
title: "Single Species Gopher Model with Seasonality"
author: "Elizabeth Hiroyasu"
date: "July 18, 2016"
output: html_document
---
The SSGophers.Rmd file contains a single species annual model for gopher populations. However, we expect that seasonality may be an important aspect determining gopher population growth. This model includes seasonality (summer and winter) in the matrix model presented previously.

Again, parameters and initial population sizes have been taken from literature estimates both of the growth and survival rates of the population and the densities of gopher populations for the species Thomomys bottae. For this model we used the 'popbio' package in R.
```{r}
library(popbio)
```

##Setting up the matrix
We directly filled population parameters into a 3x3 matrix, with a juvenile, subadult, and adult stage. 

```{r}
stages = c("Juvenile", "SubAdult", "Adult")
```

The initial matrix looks as follows, for each season:

 Stage   | Juvenile | SubAdult | Adult
---------|----------|----------|-------            
Juvenile | 0        | f2       | f3 
SubAdult | p1       | 0        | 0
Adult    | 0        | p2       | p3 

p1: probability of juvenile becoming a sub adult*
p2: probability of a subadult becoming reproductive adult
p3: probability of staying adult

f2: birth rate from subadults
f3: birth rate from reproductive adults

*this term is unknown in gophers so we will use estimates from the vole population

Seasonal Matrices:
Note that for now, fecundity is held constant because of the lack of differentiation in the data between subadult and adult breeding. P2 is a made up number for the winter matrix.
```{r}
SummerMatrix<-matrix(data=c(0,2.535, 7.775, 0.69, 0, 0, 0, 0.433, 0.777),
              nrow=3, ncol=3, byrow=TRUE,
              dimnames=list(stages,stages))

WinterMatrix<-matrix(data=c(0, 2.535, 7.775, 0.69,0,0,0, 0.33, 0.61), 
                      nrow=3, ncol=3, byrow=TRUE,dimnames = list(stages, stages))
```

Taking the matrix product to create a matrix starting in summer, we get the following:
```{r, echo=FALSE}
MatrixProduct<-SummerMatrix%*%WinterMatrix
MatrixProduct
```



Combining the matrices into a single array
```{r}
MatrixArray<-array(c(SummerMatrix, WinterMatrix), c(3,3,3))
MatrixArray
```


